Skew-symmetric entropy stable discontinuous Galerkin formulations with applications to hybrid meshes

### 2019-02-14 docs/skewform/main.tex

\date{}
\author{Jesse Chan}
\title{Skew-symmetric entropy stable discontinuous Galerkin formulations}% with applications to hybrid meshes}
\titlerunning{Skew-symmetric entropy stable DG formulations}

@@

%\tableofcontents

@@

However, when applied to nonlinear conservation laws, high order methods can experience artificial growth and blow-up near under-resolved features such as shocks, turbulence, or boundary layers.  In practice, the application of high order methods to practical problems requires shock capturing and stabilization techniques (such as artificial viscosity) or solution regularization (such as filtering or limiting) to prevent solution blow-up.  The resulting schemes for nonlinear conservation laws walk a fine line between stability, robustness, and accuracy.  Aggressive stabilization or regularization can result in the loss of high order accuracy, while too little can result in instability \cite{wang2013high}.  Moreover, it can be difficult to determine robust expressions for stabilization paramaters, as parameters which work for one simulation can fail when applied to a different physical regime or discretization setting.

@@

We are interested in addressing cases where the generalized SBP property (and as a result, the strong decoupled SBP property in Theorem~\ref{lemma:dsbp}) do not hold.

@@

\bm{1}^T\bm{W}\pd{U(\bm{u}_q)}{t} + \sum_{i=1}^d\bm{1}^T\bm{B}^i \LRp{\psi_i(\tilde{\bm{u}}_f) - \tilde{\bm{v}}_f^T\bm{f}_i^*} = 0, \qquad \bm{u}_q = \bm{V}_q\bm{u}.

@@

@@

One can show that \cite{chan2017discretely}

@@

A common approach is to rewrite the geometric terms as the curl of some quantity $\bm{r}^i$, but to interpolate $\bm{r}^i$ before applying the curl \cite{visbal2002use, kopriva2006metric, hindenlang2012explicit, chan2018discretely}:
\begin{gather}
\bm{r}^i = { \pd{\bm{x}}{\hat{x}_i}\times \bm{x}}, \qquad
\LRs{\begin{array}{c}
{G}^k_{1j}\\
{G}^k_{2j}\\
{G}^k_{3j}\end{array}} = \begin{bmatrix}
\LRp{-\hat{\Grad}\times I_{N_{\rm geo}}\LRp{x_3\hat{\Grad}x_2}}_j\\
\LRp{\hat{\Grad}\times I_{N_{\rm geo}}\LRp{x_3\hat{\Grad}x_1}}_j\\
\LRp{\hat{\Grad}\times I_{N_{\rm geo}}\LRp{x_1\hat{\Grad}x_2}}_j
\end{bmatrix},\\
%-\frac{1}{2}\LRp{\pd{I_{N_{\rm geo}}\bm{r}^k}{\hat{x}_j}-\pd{I_{N_{\rm geo}}\bm{r}^j}{\hat{x}_k}},\\
N_{\rm geo} \leq \begin{cases}
N+1 & \text{(tetrahedra)}\\
N & \text{(hexahedra)}
\end{cases},\nonumber
\label{eq:iconscurl}
\end{gather}
where $I_{N_{\rm geo}}$ denotes a degree $N_{\rm geo}$ polynomial interpolation operator with appropriate interpolation nodes.\footnote{This interpolation step must be performed using interpolation points with an appropriate number of nodes on each boundary \cite{chan2018discretely}.  These include, for example, GLL nodes on tensor product elements, as well as Warp and Blend nodes on non-tensor product elements \cite{warburton2006explicit, chan2015comparison}.}  The restriction on the maximum value of $N_{\rm geo}$ ensures that $G^k_{ij} \in V^N$ (e.g. $G^k_{ij} \in P^N$ on tetrahedral elements and $G^k_{ij}\in Q^N$ on hexahedral elements), which is also necessary to guarantee (\ref{eq:dgcl}).

@@

We can now construct and prove entropy conservation and free stream preservation for a skew-symmetric formulation on a curved element $D^k$.  Let $\bm{Q}^i_k$ be given by (\ref{eq:dncurved}), and define the curved mass matrix $\bm{M}^k = \bm{V}_q^T\bm{W}\diag{\bm{J}^k}\bm{V}_q$.  Note that $\bm{M}^k$ is positive-definite so long as $J^k$ is positive at all quadrature points.  We define the auxiliary quantities $\tilde{\bm{u}}$ 
\begin{gather*}
\bm{v}_q = \bm{v}\LRp{\bm{u}_q}, \qquad \tilde{\bm{v}} = \begin{bmatrix}
\bm{V}_q\\
\bm{V}_f
\end{bmatrix}\bm{P}^k_q\bm{v}_q, \qquad \tilde{\bm{u}} = \bm{u}\LRp{\tilde{\bm{v}}}.
\end{gather*}
where $\bm{P}^k_q = \LRp{\bm{M}^k}^{-1}\bm{V}_q^T\bm{W}\diag{\bm{J}^k}$.  Then, we have the following theorem:
\begin{theorem}
\label{thm:skewformcurved}
Let the geometric terms $G^k_{ij}$ be computed from (\ref{eq:iconscurl}), let the scaled outward normals $\bm{n}^k_i$ be computed pointwise from (\ref{eq:normals}), and suppose Assumption~\ref{ass:quad} holds for $v = 1$ and $v= G^k_{ij}$.  Let $\tilde{\bm{u}}_f^+$ denote the face value of the entropy-projected conservative variables $\tilde{\bm{u}}_f$ on the neighboring element.  Then, the formulation
\begin{gather}
\bm{M}^k\pd{\bm{u}_h}{t} + 
\sum_{i=1}^d \LRs{\begin{array}{cc}
\bm{V}_q \\
\bm{V}_f\end{array}}^T \LRp{\LRp{\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T} \circ \bm{F}^i_S}\bm{1} + \bm{V}_f^T \bm{W}_f \diag{{\bm{n}^k_i}}\bm{f}_i^* = 0, \label{eq:skewformcurved}\\
%\sum_{i=1}^d \LRs{\begin{array}{cc}
%\bm{P}_q & \bm{L}_f\end{array}} \LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{L}_f \diag{{\bm{n}_i}}\bm{f}_i^* = 0, \label{eq:skewform}\\
\LRp{\bm{F}^i_S}_{ij} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q, \nonumber\\
\bm{f}_i^* = \bm{f}^i_S(\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f), \qquad \text{ on interior interfaces,} \nonumber
\end{gather}
is semi-discretely entropy conservative on $D^k$ such that for $\bm{u}_q = \bm{V}_q\bm{u}$,
\begin{gather*}
\bm{1}^T\bm{W}\diag{\bm{J}^k}\pd{U(\bm{u}_q)}{t} + \sum_{i=1}^d\bm{1}^T\bm{W}_f \diag{\bm{n}^k_i} \LRp{\psi_i(\tilde{\bm{u}}_f) - \tilde{\bm{v}}_f^T\bm{f}_i^*} = 0.
\end{gather*}
Additionally, the method is free-stream preserving such that $\pd{\bm{u}_h}{t} = 0$ for constant solutions.
\end{theorem}
\begin{proof}
The proof of entropy stability is identical to that of Theorem~\ref{thm:esdg}, except that Lemma~\ref{lemma:vdsbpcurved} is used instead of Lemma~\ref{lemma:sbpcor}.  However, the proof of free-stream preservation differs from proofs found in the literature, which typically show that the right hand side is zero for $\bm{u}_h$ constant \cite{crean2018entropy, chan2018discretely}.  This strategy for is appropriate for ``strong'' formulations, but must be modified for the skew-symmetric formulation (\ref{eq:skewformcurved}).

@@

\subsection{On quadrature conditions for Assumption~\ref{ass:quad} for $v = 1$ and $v = G^k_{ij}$}
\label{sec:curvgeo}
Semi-discrete entropy conservation on curved meshes requires that Assumption~\ref{ass:quad} holds for $v = 1$ and $v = G^k_{ij}$.  We discuss specific choices of volume and surface quadrature on for which this assumption is valid, and summarize the maximum degree $N_{\rm geo}$ of the polynomial geometric approximation under which entropy stability holds for common choices of volume and surface quadrature.  
%\note{Emphasize we're interested in taking $N_{\rm geo}$ as large as possible for a given quadrature rule, and that there is a focus on degree $(2N-1)$ quadrature rules.}

\subsubsection{Simplicial elements}

On simplicial elements, for $u \in P^N$, $\pd{u}{\hat{x}_j}\in P^{N-1}$, and for $N_{\rm geo}\leq (N+1)$, $G^k_{ij}\in P^{N_{\rm geo}-1}$.  Thus, the integrands in Assumption~\ref{ass:quad} $\pd{u}{\hat{x}_j} v \in P^{N+N_{\rm geo}-1}$ and $uv n_i \in P^{N+N_{\rm geo}}$ for $v = G^k_{ij}$.  A volume quadrature rule of degree $(2N-1)$ and surface quadrature of degree $2N$ are then sufficient to ensure semi-discrete entropy conservation for $N_{\rm geo}\leq N$.  However, that in order to ensure that the mass matrix is positive-definite in Assumption~\ref{ass:quad}, the volume quadrature must be degree $2N$ in general on simplices.  If the volume quadrature is instead exact for degree $2N$ polynomials, then entropy conservation is guaranteed up to $N_{\rm geo} \leq N+1$.  
%\note{Add discussion on triangular elements.}

\subsubsection{Tensor product elements}

%Tensor product $(N+1)$-point GLL quadratures are a common choice for quadrilateral and hexahedral elements, and are exact for polynomials of degree $(2N-1)$ in each coordinate direction.  
It was shown in Section~\ref{sec:assump1} that tensor product quadratures of degree $(2N-1)$ satisfy Assumption~\ref{ass:quad} holds for $v = 1$.  However, in contrast to the simplicial case, it is not immediately clear that degree $(2N-1)$ volume quadratures exactly integrate $\int_{\hat{D}} \pd{u}{\hat{x}_j}v$ for $v = G^k_{ij}$ for tensor product elements.  The difference between simplicial and tensor product elements is the polynomial space in which the derivative lies.  In contrast to the simplicial case, if $u \in Q^N$, $\pd{u}{\hat{x}_j} \not\in Q^{N-1}$.  Consider the three-dimensional case with $u, v \in Q^N$ and $i = 1$.  Then, differentiation reduces the polynomial degree in one coordinate but not others and $\pd{u}{\hat{x}_1} \in Q^{N-1,N,N}$.  As a result, $\pd{u}{\hat{x}_j}v \not\in Q^{2N-1}$, and a tensor product quadrature of degree $(2N-1)$ does not exactly integrate $\int_{\hat{D}} \pd{u}{\hat{x}_j}v$ for general $v\in Q^N$.

We first consider the two dimensional case with volume quadrature of degree $(2N-1)$ and show that Assumption~\ref{ass:quad} holds with $v = G^k_{ij}$ for all $N_{\rm geo} \leq N$.  On quadrilateral elements, $G^k_{ij}$ is
\begin{align*}
G^k_{11} &= \pd{x_2}{\hat{x}_2} \in Q^{N_{\rm geo}, N_{\rm geo}-1}, \qquad G^k_{12} = -\pd{x_2}{\hat{x}_1} \in Q^{N_{\rm geo}-1, N_{\rm geo}},\\
G^k_{21} &= -\pd{x_1}{\hat{x}_2} \in Q^{N_{\rm geo}, N_{\rm geo}-1}, \qquad G^k_{22} = \pd{x_1}{\hat{x}_1}\in Q^{N_{\rm geo}-1, N_{\rm geo}}.
\end{align*}
Since $\pd{u}{\hat{x}_1} \in Q^{N-1,N}$ and $\pd{u}{\hat{x}_2} \in Q^{N,N-1}$
\[
\pd{u}{\hat{x}_i}G^k_{ij} \in Q^{N+N_{\rm geo}-1} \subset Q^{2N-1}, \qquad \forall N_{\rm geo}\leq N.
\]
Thus, any quadrature rule of degree $(2N-1)$ exactly integrates $\int_{\hat{D}}\pd{u}{\hat{x}_i}v$ for $v = G^k_{ij}$.

We now consider the condition in Assumption~\ref{ass:quad} on the surface integrals $\int_{\partial \hat{D}} u v \hat{n}_j$ for $v= G^k_{ij}$.  For left and right faces of the quadrilateral, $\hat{n}_2 = 0$, so this condition reduces to ensuring that the quantity $u G^k_{i1}$ is integrated exactly using quadrature for $i = 1,2$.  Since $G^k_{i1}$ are degree $N_{\rm geo}-1$ in the $\hat{x}_2$ coordinate, $G^k_{i1}$ is degree $N_{\rm geo}-1$ and $uG^k_{i1} \hat{n}_1 \in Q^{N+N_{\rm geo}-1}$ along the left and right faces.  Similarly, $uG^k_{i2} \hat{n}_2 \in Q^{N+N_{\rm geo}-1}$ along the top and bottom faces and are zero along the left and right faces.  Thus, any surface quadrature rule exact for degree $N+N_{\rm geo}-1$ polynomials satisfies Assumption~\ref{ass:quad}, and for a degree $(2N-1)$ surface quadrature, entropy stability is guaranteed for any $N_{\rm geo}\leq N$.

We now consider the three-dimensional case.  Unlike quadrilateral elements, Assumption~\ref{ass:quad} does not hold under degree $(2N-1)$ volume quadrature for $N_{\rm geo} = N$ on general curved hexahedra.  Expanding out (\ref{eq:iconscurl}) gives
\[
G^k_{11} = \pd{}{\hat{x}_3} I_{N_{\rm geo}}\LRp{{x}_3 \pd{x_2}{\hat{x}_2}} - \pd{}{\hat{x}_2} I_{N_{\rm geo}}\LRp{{x}_3 \pd{x_2}{\hat{x}_3}} \in Q^{N_{\rm geo}}.
\]
Repeating for the other geometric terms, one can show that $G^k_{ij} \in Q^{N_{\rm geo}}$ on hexahedral elements.  Thus, if $u\in Q^N$, $\pd{u}{\hat{x}_1} G^k_{i1} \in Q^{N+N_{\rm geo}-1,N+N_{\rm geo},N+N_{\rm geo}}$, and is only integrated exactly by volume quadratures of degree $(2N-1)$ for geometric degrees $N_{\rm geo}\leq (N-1)$.  Similarly, Assumption~\ref{ass:quad} does not hold under degree $(2N-1)$ surface quadratures unless $N_{\rm geo} \leq (N-1)$, due to the fact that traces of $G^k_{ij}$ are degree $N_{\rm geo}$ polynomials in each coordinate.\footnote{It is possible to  construct the geometric terms for $N_{\rm geo} = N$ using a local $H_{\rm div}$ basis where 
\[
\bm{r}^i \in Q^{N-1,N,N} \times Q^{N,N-1,N} \times Q^{N,N,N-1}.
\]  
Then, the geometric terms satisfy $\Grad \times \bm{r}^i \in Q^{N,N-1,N-1}\times Q^{N-1,N,N-1} \times Q^{N-1,N-1,N}$ with traces in $Q^{N-1}$, and Assumption~\ref{ass:quad} holds under degree $(2N-1)$ volume and surface quadrature.  This approach will be investigated in more detail in future work.  %Numerical results show that this approach produces geometric terms which are up to an order of magnitude more accurate than those constructed simply by taking $N_{\rm geo} \leq (N-1)$.  This approach will be investigated in more detail in future work.
}

Most implementations on tensor product elements utilize volume and surface quadratures of either degree $(2N-1)$ or $2N$.  We summarize below for different pairings of volume and surface quadrature the maximum degree $N_{\rm geo}$ under which Assumption~\ref{ass:quad} is satisfied and entropy stability is guaranteed:
\begin{enumerate}
\item On quadrilateral elements, Assumption~\ref{ass:quad} holds for $N_{\rm geo} = N$ and any tensor product volume and surface quadratures of degree $(2N-1)$ 
\item On hexahedral elements, Assumption~\ref{ass:quad} holds for $N_{\rm geo} = N-1$ and any tensor product volume and surface quadratures of degree $(2N-1)$.  If the SBP property holds (for example, when the volume and surface quadratures are of degree $2N$) then Assumption~\ref{ass:quad} holds for $N_{\rm geo} = N$.
\end{enumerate}

We note that the condition $N_{\rm geo} = N-1$ is non-standard for tensor product elements.  However, this condition is only necessary for entropy stability when $\bm{Q}^i$ does not satisfy a generalized SBP property (see Remark~\ref{remark:dsbp}).  To the author's knowledge, this setting has not been considered within the literature on tensor product elements.

%\note{Finish}
%\begin{remark}
%We note that this work assumes that the same quadrature is used to volume contributions in each direction, and can be refined slightly 
%\note{add line DG}

>>>>

\date{}
\author{Jesse Chan}
\title{Skew-symmetric entropy stable discontinuous Galerkin formulations with applications to hybrid meshes}
\titlerunning{Skew-symmetric entropy stable DG formulations}

@@

\tableofcontents

@@

However, when applied to nonlinear conservation laws, high order methods can experience artificial growth and blow-up near under-resolved features such as shocks or turbulence.  In practice, the application of high order methods to practical problems requires shock capturing and stabilization techniques (such as artificial viscosity) or solution regularization (such as filtering or limiting) to prevent solution blow-up.  The resulting schemes for nonlinear conservation laws walk a fine line between stability, robustness, and accuracy.  Aggressive stabilization or regularization can result in the loss of high order accuracy, while too little can result in instability \cite{wang2013high}.  Moreover, it can be difficult to determine robust expressions for stabilization paramaters, as parameters which work for one simulation can fail when applied to a different physical regime or discretization setting.

@@

When a generalized SBP property holds for $\bm{Q}^i$, entropy stability can be proven using the strong SBP property in Theorem~\ref{lemma:dsbp} \cite{chan2017discretely, chan2018discretely}.  The focus of this work is to address cases where the generalized SBP property (and as a result, the SBP property in Theorem~\ref{lemma:dsbp}) do not hold.

@@

\bm{1}^T\bm{W}\pd{U(\bm{u}_q)}{t} + \sum_{i=1}^d\bm{1}^T\bm{B}^i \LRp{\tilde{\bm{v}}_f^T\bm{f}_i^* - \psi_i(\tilde{\bm{u}}_f)} = 0, \qquad \bm{u}_q = \bm{V}_q\bm{u}.

@@

\label{eq:entropyrhs}

@@

Algebraic manipulations (see \cite{chan2017discretely} for details) yields that 

@@

To ensure that the geometric terms satisfy the GCL, we first rewrite the geometric terms as the curl of some quantity $\bm{r}^i$, but interpolate $\bm{r}^i$ before applying the curl \cite{visbal2002use, kopriva2006metric, hindenlang2012explicit, chan2018discretely}:
\begin{gather}
\bm{r}^i = { \pd{\bm{x}}{\hat{x}_i}\times \bm{x}}, \qquad
\LRs{\begin{array}{c}
{G}^k_{1j}\\
{G}^k_{2j}\\
{G}^k_{3j}\end{array}} = \begin{bmatrix}
\LRp{-\hat{\Grad}\times I_{N_{\rm geo}}\LRp{x_3\hat{\Grad}x_2}}_j\\
\LRp{\hat{\Grad}\times I_{N_{\rm geo}}\LRp{x_3\hat{\Grad}x_1}}_j\\
\LRp{\hat{\Grad}\times I_{N_{\rm geo}}\LRp{x_1\hat{\Grad}x_2}}_j
\end{bmatrix},\\
%-\frac{1}{2}\LRp{\pd{I_{N_{\rm geo}}\bm{r}^k}{\hat{x}_j}-\pd{I_{N_{\rm geo}}\bm{r}^j}{\hat{x}_k}},\\
N_{\rm geo} \leq \begin{cases}
N+1 & \text{(tetrahedra)}\\
N & \text{(hexahedra)}
\end{cases},\nonumber
\label{eq:iconscurl}
\end{gather}
where $I_{N_{\rm geo}}$ denotes a degree $N_{\rm geo}$ polynomial interpolation operator with appropriate interpolation nodes.\footnote{This interpolation step must be performed using interpolation points with an appropriate number of nodes on each boundary \cite{chan2018discretely}.  These include, for example, GLL nodes on tensor product elements, and optimized interpolation nodes on non-tensor product elements \cite{hesthaven1998electrostatics, warburton2006explicit, chan2015comparison}.}  The restriction on the maximum value of $N_{\rm geo}$ ensures that $G^k_{ij} \in V^N$ (e.g. $G^k_{ij} \in P^N$ on tetrahedral elements and $G^k_{ij}\in Q^N$ on hexahedral elements), which is also necessary to guarantee (\ref{eq:dgcl}).

@@

We can now construct and prove entropy conservation and free stream preservation for a skew-symmetric formulation on a curved element $D^k$.  Let $\bm{Q}^i_k$ be given by (\ref{eq:dncurved}), and define the curved mass matrix $\bm{M}^k = \bm{V}_q^T\bm{W}\diag{\bm{J}^k}\bm{V}_q$.  Note that $\bm{M}^k$ is positive-definite so long as $J^k$ is positive at all quadrature points.  We define the auxiliary quantities $\tilde{\bm{u}}$ 
\begin{gather*}
\bm{v}_q = \bm{v}\LRp{\bm{u}_q}, \qquad \tilde{\bm{v}} = \begin{bmatrix}
\bm{V}_q\\
\bm{V}_f
\end{bmatrix}\bm{P}^k_q\bm{v}_q, \qquad \tilde{\bm{u}} = \bm{u}\LRp{\tilde{\bm{v}}}.
\end{gather*}
where $\bm{P}^k_q = \LRp{\bm{M}^k}^{-1}\bm{V}_q^T\bm{W}\diag{\bm{J}^k}$.  Then, we have the following theorem:
\begin{theorem}
\label{thm:skewformcurved}
Let the geometric terms $G^k_{ij}$ be computed from (\ref{eq:iconscurl}), let the scaled outward normals $\bm{n}^k_i$ be computed pointwise from (\ref{eq:normals}), and suppose Assumption~\ref{ass:quad} holds for $v = 1$ and $v= G^k_{ij}$.  Let $\tilde{\bm{u}}_f^+$ denote the face value of the entropy-projected conservative variables $\tilde{\bm{u}}_f$ on the neighboring element.  Then, the formulation
\begin{gather}
\bm{M}^k\pd{\bm{u}_h}{t} + 
\sum_{i=1}^d \LRs{\begin{array}{cc}
\bm{V}_q \\
\bm{V}_f\end{array}}^T \LRp{\LRp{\bm{Q}^i_k - \LRp{\bm{Q}^i_k}^T} \circ \bm{F}^i_S}\bm{1} + \bm{V}_f^T \bm{W}_f \diag{{\bm{n}^k_i}}\bm{f}_i^* = 0, \label{eq:skewformcurved}\\
%\sum_{i=1}^d \LRs{\begin{array}{cc}
%\bm{P}_q & \bm{L}_f\end{array}} \LRp{\LRp{\bm{D}^i_N - \bm{W}_N^{-1}\LRp{\bm{Q}^i_N}^T} \circ \bm{F}_S}\bm{1} + \bm{L}_f \diag{{\bm{n}_i}}\bm{f}_i^* = 0, \label{eq:skewform}\\
\LRp{\bm{F}^i_S}_{ij} = \bm{f}^i_S\LRp{\tilde{\bm{u}}_i,\tilde{\bm{u}}_j}, \qquad 1\leq i,j\leq N_q + N^f_q, \nonumber\\
\bm{f}_i^* = \bm{f}^i_S(\tilde{\bm{u}}_f^+,\tilde{\bm{u}}_f), \qquad \text{ on interior interfaces,} \nonumber
\end{gather}
is semi-discretely entropy conservative on $D^k$ such that for $\bm{u}_q = \bm{V}_q\bm{u}$,
\begin{gather*}
\bm{1}^T\bm{W}\diag{\bm{J}^k}\pd{U(\bm{u}_q)}{t} + \sum_{i=1}^d\bm{1}^T\bm{W}_f \diag{\bm{n}^k_i} \LRp{\psi_i(\tilde{\bm{u}}_f) - \tilde{\bm{v}}_f^T\bm{f}_i^*} = 0.
\end{gather*}
Additionally, the method is free-stream preserving such that $\pd{\bm{u}_h}{t} = 0$ for constant solutions.
\end{theorem}
\begin{proof}
The proof of entropy stability is identical to that of Theorem~\ref{thm:esdg}, except that Lemma~\ref{lemma:vdsbpcurved} is used instead of Lemma~\ref{lemma:sbpcor}.  However, proving free-stream preservation differs from approaches in the literature, which typically show that the right hand side is zero for $\bm{u}_h$ constant \cite{crean2018entropy, chan2018discretely}.  This strategy for is appropriate for ``strong'' formulations, but must be modified for the skew-symmetric formulation (\ref{eq:skewformcurved}).

@@

\section{Conditions on the accuracy of volume and surface quadrature rules}

The previous sections outline minimal conditions under which entropy stability is guaranteed under a skew-symmetric formulation and a polynomial geometric mapping.  In Section~\ref{sec:curvgeo}, we translate these minimial conditions into conditions on quadrature accuracy.

Additionally, while it is possible to maintain entropy stability under reduced accuracy quadratures, the accuracy of the resulting schemes depend on the accuracy of the volume and surface quadrature rules.  We analyze the degree of accuracy expected based on the accuracy of different quadrature rules in Section~\ref{sec:accskew}.

\subsection{On quadrature conditions for Assumption~\ref{ass:quad} for $v = 1$ and $v = G^k_{ij}$}
\label{sec:curvgeo}

Semi-discrete entropy conservation on curved meshes requires that Assumption~\ref{ass:quad} holds for $v = 1$ and $v = G^k_{ij}$.  We discuss specific choices of volume and surface quadrature on for which this assumption is valid, and summarize the maximum degree $N_{\rm geo}$ of the polynomial geometric approximation under which entropy stability holds for common choices of volume and surface quadrature.  
%\note{Emphasize we're interested in taking $N_{\rm geo}$ as large as possible for a given quadrature rule, and that there is a focus on degree $(2N-1)$ quadrature rules.}

In order to ensure that the mass matrix is positive-definite in Assumption~\ref{ass:quad}, the volume quadrature must be degree $2N$ in general on simplices.  The following lemma summarizes expected behavior for surface quadrature rules of varying order:
\begin{lemma}
Let $\hat{D}$ be a simplex with volume quadrature which is exact for degree $2N$ polynomials.  Let the surface quadrature be exact for polynomials of degree $M+N$.  Then, the skew-symmetric formulation (\ref{eq:skewformcurved}) is entropy stable for $N_{\rm geo} \leq \min\LRp{N+1,M+1}$.
\label{lemma:curvsimp}
\end{lemma}
\begin{proof}
Entropy stability holds if Assumption~\ref{ass:quad} holds for $v = 1$ and $v = G^k_{ij}$.  
Simplicial elements require $N_{\rm geo}\leq (N+1)$ in order to guarantee that $G^k_{ij}\in P^{N_{\rm geo}-1} \subset P^{N}$, which is necessary to satisfy the discrete GCL \cite{chan2018discretely}.  Then, for $u \in P^N$, $\pd{u}{\hat{x}_j}\in P^{N-1}$, the integrands in Assumption~\ref{ass:quad} are $\pd{u}{\hat{x}_j} v \in P^{N+N_{\rm geo}-2}$ and $uv n_i \in P^{N+N_{\rm geo}-1}$ for $v = G^k_{ij}$.  The volume quadrature exactly integrates the first integrand for $N_{\rm geo} \leq N+1$, while the surface quadrature exactly integrates the second integrand for $M \geq N_{\rm geo}-1$, or $N_{\rm geo} \leq M+1$.
\end{proof}

The situation is more complicated for curved tensor product elements.  It was shown in Section~\ref{sec:assump1} that tensor product quadratures of degree $(2N-1)$ satisfy Assumption~\ref{ass:quad} holds for $v = 1$.  However, in contrast to the simplicial case, it is not immediately clear that degree $(2N-1)$ volume quadratures exactly integrate $\int_{\hat{D}} \pd{u}{\hat{x}_j}v$ for $v = G^k_{ij}$ for tensor product elements.  The difference between simplicial and tensor product elements is the polynomial space in which the derivative lies.  In contrast to the simplicial case, if $u \in Q^N$, $\pd{u}{\hat{x}_j} \not\in Q^{N-1}$.  Consider the three-dimensional case with $u, v \in Q^N$ and $i = 1$.  Then, differentiation reduces the polynomial degree in one coordinate but not others and $\pd{u}{\hat{x}_1} \in Q^{N-1,N,N}$.  As a result, $\pd{u}{\hat{x}_j}v \not\in Q^{2N-1}$, and a tensor product quadrature of degree $(2N-1)$ (in each coordinate) does not exactly integrate $\int_{\hat{D}} \pd{u}{\hat{x}_j}v$ for general $v\in Q^N$.

We address the quadrilateral case first:
\begin{lemma}
Let $\hat{D}$ be a quadrilateral.  Suppose the volume quadrature be exact for degree $M+N$ polynomials, and that the surface quadrature be exact for polynomials of degree $M+N$.  Then, the skew-symmetric formulation (\ref{eq:skewformcurved}) is entropy stable for $N_{\rm geo} \leq \min\LRp{N,M+1}$.
\label{lemma:curvquad}
\end{lemma}
\begin{proof}
As in Lemma~\ref{lemma:curvsimp}, entropy stability holds if Assumption~\ref{ass:quad} holds for $v = 1$ and $v = G^k_{ij}$.  The case of $v = 1$ was addressed previously, and we focus on $v = G^k_{ij}$.  We first characterize the polynomial degree of the geometric terms $G^k_{ij}$.  In contrast to the simplicial case, tensor product elements require $N_{\rm geo}\leq N$ in order to ensure that $G^k_{ij}\in Q^{N,N}$ and that the discrete GCL is satisfied \cite{kopriva2006metric}.  On a quadrilateral element with a degree $N_{\rm geo}$ geometric mapping, $G^k_{ij}$ is
% with volume quadrature of degree $(2N-1)$ and show that Assumption~\ref{ass:quad} holds with $v = G^k_{ij}$ for all $N_{\rm geo} \leq N$.  
\begin{align*}
G^k_{11} &= \pd{x_2}{\hat{x}_2} \in Q^{N_{\rm geo}, N_{\rm geo}-1}, \qquad G^k_{12} = -\pd{x_2}{\hat{x}_1} \in Q^{N_{\rm geo}-1, N_{\rm geo}},\\
G^k_{21} &= -\pd{x_1}{\hat{x}_2} \in Q^{N_{\rm geo}, N_{\rm geo}-1}, \qquad G^k_{22} = \pd{x_1}{\hat{x}_1}\in Q^{N_{\rm geo}-1, N_{\rm geo}}.\nonumber
\end{align*}
Since $\pd{u}{\hat{x}_1} \in Q^{N-1,N}$ and $\pd{u}{\hat{x}_2} \in Q^{N,N-1}$
\[
\pd{u}{\hat{x}_i}G^k_{ij} \in Q^{N+N_{\rm geo}-1}.% \subset Q^{2N-1}, \qquad \forall N_{\rm geo}\leq N.
\]
The volume quadrature exactly integrates this integrand for $M \geq N_{\rm geo}-1$.

We now consider the condition in Assumption~\ref{ass:quad} on the surface integrals $\int_{\partial \hat{D}} u v \hat{n}_j$ for $v= G^k_{ij}$.  For left and right faces of the quadrilateral, $\hat{n}_2 = 0$, so this condition reduces to ensuring that the quantity $u G^k_{i1}$ is integrated exactly using quadrature for $i = 1,2$.  Since $G^k_{i1}$ are degree $N_{\rm geo}-1$ in the $\hat{x}_2$ coordinate, $G^k_{i1}$ is degree $N_{\rm geo}-1$ and $uG^k_{i1} \hat{n}_1 \in Q^{N+N_{\rm geo}-1}$ along the left and right faces.  Similarly, $uG^k_{i2} \hat{n}_2 \in Q^{N+N_{\rm geo}-1}$ along the top and bottom faces and are zero along the left and right faces.  The surface quadrature rule exactly integrates such integrands for $M \geq N_{\rm geo} - 1$, or $N_{\rm geo} \leq M+1$.
\end{proof}

Existing proofs of entropy stability on quadrilaterals rely on $(N+1)$-point GLL volume and surface quadratures, which are exact for degree $2N-1$ polynomials.   The novelty of Lemma~\ref{lemma:curvquad} is that the proof holds for any combination of degree $2N-1$ volume and surface quadratures (for example, $(N+1)$-point GLL volume quadrature and an $(N-1)$-point Gauss surface quadrature).

We now consider the three-dimensional case.  In contrast to the quadrilateral case, the GCL is not automatically satisfied for a degree $N_{\rm geo} \leq N$ geometric mapping.  Instead, GCL-satisfying geometric terms are approximated using (\ref{eq:iconscurl}).  Expanding out the expression for $G^k_{11}$ gives
\[
G^k_{11} = \pd{}{\hat{x}_3} I_{N_{\rm geo}}\LRp{{x}_3 \pd{x_2}{\hat{x}_2}} - \pd{}{\hat{x}_2} I_{N_{\rm geo}}\LRp{{x}_3 \pd{x_2}{\hat{x}_3}} \in Q^{N_{\rm geo}}.
\]
Repeating for the other geometric terms, one can show that $G^k_{ij} \in Q^{N_{\rm geo}}$ on hexahedral elements.  Thus, if $u\in Q^N$, $\pd{u}{\hat{x}_1} G^k_{i1} \in Q^{N+N_{\rm geo}-1,N+N_{\rm geo},N+N_{\rm geo}}$, and is only integrated exactly by volume quadratures of degree $(2N-1)$ for geometric degrees $N_{\rm geo}\leq (N-1)$.  Similarly, Assumption~\ref{ass:quad} does not hold under degree $(2N-1)$ surface quadratures unless $N_{\rm geo} \leq (N-1)$, due to the fact that traces of $G^k_{ij}$ are degree $N_{\rm geo}$ polynomials in each coordinate.\footnote{It is possible to  construct the geometric terms for $N_{\rm geo} = N$ using a local $H_{\rm div}$ basis where 
\[
\bm{r}^i \in Q^{N-1,N,N} \times Q^{N,N-1,N} \times Q^{N,N,N-1}.
\]  
Then, the geometric terms satisfy $\Grad \times \bm{r}^i \in Q^{N,N-1,N-1}\times Q^{N-1,N,N-1} \times Q^{N-1,N-1,N}$ with traces in $Q^{N-1}$, and Assumption~\ref{ass:quad} holds under degree $(2N-1)$ volume and surface quadrature.  This approach will be investigated in more detail in future work.  %Numerical results show that this approach produces geometric terms which are up to an order of magnitude more accurate than those constructed simply by taking $N_{\rm geo} \leq (N-1)$.  This approach will be investigated in more detail in future work.
}
We summarize these findings in the following lemma for hexahedral elements:
\begin{lemma}
Let $\hat{D}$ be a hexahedral element, with geometric terms constructed using (\ref{eq:iconscurl}).  Let the volume quadrature be exact for degree $M+N$ polynomials, and let the surface quadrature be exact for polynomials of degree $M+N$.  Then, the skew-symmetric formulation (\ref{eq:skewformcurved}) is entropy stable for $N_{\rm geo} \leq \min\LRp{N,M}$.
\label{lemma:curvquad}
\end{lemma}
